---
title: "cen_fsus group project"
output: html_document
---
#Intro information
A	Higher managerial, administrative and professional	
B	Intermediate managerial, administrative and professional	
C1	Supervisory, clerical and junior managerial, administrative and professional	
C2	Skilled manual workers
D	Semi-skilled and unskilled manual workers 
E	State pensioners, casual and lowest grade workers, unemployed with state benefits only

The census data calssified the social grades per groups. AB, C1, C2, DE. IT is exoected to hav

```{r setup, include=FALSE}
library(readr)
census <- read_csv("C:/Users/helen/Desktop/Imperial/Math_Stat/Group_project_R/filtered data.csv")
View(census)
library(dplyr)
census$region <- factor(census$region, levels = c(1, 2, 3, 4, 5, 6, 7,
                                                       8, 9, 10, 11), 
                           labels = c('North East', 'North West', 'Yorkshire and the Humber', 
                                      'East Midlands', 'West Midlands', 'East of England', 
                                      'Inner London',
                                      'Outer London', 'South East', 'South West', 'Wales'))
cen <- select(census, caseno, marstat, hlqupuk11, scgpuk11c)
cen$separated <- cen$marstat
cen <- cen %>% mutate(separated = replace(separated, marstat == 2 | marstat == 3, 0))
cen <- cen %>% mutate(separated = replace(separated, marstat == 5 | marstat == 4, 1))
colnames(cen) <- c("ID", "marstat", "education", "social_grade", "separated")
#replace -9 with NA 
cen_f <- cen %>% mutate_all(funs(replace(., . == -9, NA)))
cen_f <- cen_f %>% filter(!is.na(social_grade), !is.na(education), !(education == 16))
dim(cen_f)
```

For education level "other" is a foreign degree and therefore not applicable and excluded. Attenteion when we look at immigration backround, as those people might be crossing. 

Social grade: 
1 AB (highest)
2 C1
3 C2
4 DE (lowest)
-9 were excluded as it represents "persons resident in a communal establishment OR persons aged under 16 OR persons aged over 64 OR student/schoolchild living away during term-time OR short-term migrant" (Attention when we match with other data which might include people over 64)

After filtering out people whose education level cannot be interpretated and filtering out missing values 1.684.636 observations are left. 

#quick overview
```{r}
#quick look at distributions
table(cen_f$separated)
addmargins(prop.table(table(cen_f$separated)))
table(cen_f$education)
table(cen_f$social_grade)
```
```{r}
#correlate independend variables 
library(ggplot2)
library(dplyr)
social_plot <- ggplot(cen_f, aes(x = factor(education), y = social_grade))
social_plot +
  geom_histogram(stat = "summary", fun.y = "mean", fill = "cadetblue3") +
  xlab ("Education level") +
  ylab("Average social Grade (1:highest 4: lowest)") +
  scale_x_discrete(labels = c("10" = "none", "11" = "O levels / GCSE", "12" = "CSE / GCSE grade 1 / A", "13" = "apprenticeship", "14" = "A levels", "15" = "university degree")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


cen_f %>% group_by(education) %>% summarise(n = n(), mean_grade = mean(social_grade, na.rm = TRUE))

#Spearmans ran correlation coefficient (cor between two comparative variables)
cor(cen_f$social_grade, cen_f$education, method = "spearman", use="p")

```

The data shows, that the higher the education, the higher the social grade. Exept for people who did an apprenticeship. they have a slighltly lower social grade than people with CSE/GCSE. 
There is a middle strong linear correlation between social graed and education level. The higher the education level the higher the social grade. 


```{r}
educ <- cen_f %>% group_by(education) %>% summarise(n = n(), separated = sum(separated), rel_div = (separated / n)*100)

#plot distribution of divorce per education level
#rel_div: percen_f_ft of divorce per education group
edu_plot <- ggplot(educ, aes(x = factor(education)))
edu_plot + 
  geom_col(aes(y = rel_div), fill = "cadetblue3") +
  geom_blank() +
  scale_x_discrete(labels = c("10" = "none", "11" = "O levels / GCSE", "12" = "CSE / GCSE grade 1 / A", "13" = "apprenticeship", "14" = "A levels", "15" = "university degree")) +
  scale_y_continuous(limits = c(0, 30), breaks = seq(0, 30, 3)) +
  ggtitle("Divorce rate per education level") +
  xlab ("Education level") +
  ylab ("Divore rate in %") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#Divorce rate per social grade divided by education 
```{r}
educ <- cen_f %>% group_by(education, social_grade) %>% summarise(n = n(), separated = sum(separated), rel_div = (separated / n)*100) %>% mutate(sg = n/sum(n))

#add seperation per social grade 

edu_plot <- ggplot(educ, aes(x = factor(education), y = rel_div, fill = factor(social_grade)))
edu_plot + 
  geom_col(stat = "identity", position = "dodge") +
  scale_x_discrete(labels = c("10" = "none", "11" = "O levels / GCSE", "12" = "CSE / GCSE grade 1 / A", "13" = "apprenticeship", "14" = "A levels", "15" = "university degree")) +
  scale_y_continuous(limits = c(0, 35), breaks = seq(0, 35, 3)) +
  ggtitle("Divorce rate per education level") +
  xlab ("Education level") +
  ylab ("Divore rate in %") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#plot with facet_grid per social grade 
####
edu_plot <- ggplot(educ, aes(x = factor(education), y = rel_div, fill = factor(social_grade)))
edu_plot + 
  geom_col(stat = "identity") +
  scale_x_discrete(labels = c("10" = "none", "11" = "O levels / GCSE", "12" = "CSE / GCSE grade 1 / A", "13" = "apprenticeship", "14" = "A levels", "15" = "university degree")) +
  scale_y_continuous(limits = c(0, 35), breaks = seq(0, 35, 2.5)) +
  ggtitle("Divorce rate per education level and social grade") +
  xlab ("Education level") +
  ylab ("Divore rate in %") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(~social_grade)
  
```


```{r}
soclass <- cen_f %>% filter(!is.na(social_grade)) %>% group_by(social_grade) %>% summarise(n = n(), separated = sum(separated), rel_div = (separated / n)*100)

soclass_plot <- ggplot(soclass, aes(x = factor(social_grade), y = rel_div))
soclass_plot  +
  geom_col(fill  = "cadetblue3") +
  xlab("Social grade") +
  ylab("Divorce rate in percent") +
  scale_x_discrete(labels = c("1" = "AB", "2" = "C1", "3" = "C2", "4" = "DE")) +
  scale_y_continuous(limits = c(0, 33), breaks = seq(0, 33, 5))
```

```{r}
#cen_r: data set with regions to cross check 
library(dplyr)
cen_r <- select(census, caseno, marstat, hlqupuk11, scgpuk11c, region)

cen_r$separated <- cen_r$marstat
cen_r <- cen_r %>% mutate(separated = replace(separated, marstat == 2 | marstat == 3, 0))
cen_r <- cen_r %>% mutate(separated = replace(separated, marstat == 5 | marstat == 4, 1))
colnames(cen_r) <- c("ID", "marstat", "education", "social_grade", "region", "separated")
#replace -9 with NA 
cen_r <- cen_r %>% mutate_all(funs(replace(., . == -9, NA)))
cen_r <- cen_r %>% filter(!is.na(social_grade), !is.na(education), !is.na(region), !(education == 16))
dim(cen_f)

```

#Social grade education and region

```{r}

region <- cen_r %>% group_by(region, education, social_grade) %>% summarise(n = n(), separated = sum(separated), rel_div = (separated / n)*100) %>% mutate(sg = n/sum(n))


rp <- ggplot(region, aes(x = factor(education), y = rel_div, fill = factor(social_grade)))
rp + 
  geom_col(stat = "identity", position = "dodge") +
  scale_x_discrete(labels = c("10" = "none", "11" = "O levels / GCSE", "12" = "CSE / GCSE grade 1 / A", "13" = "apprenticeship", "14" = "A levels", "15" = "university degree")) +
  scale_y_continuous(limits = c(0, 45), breaks = seq(0, 45, 3)) +
  ggtitle("Divorce rate per education level and region") +
  xlab ("Education level") +
  ylab ("Divore rate in %") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ region)
```

#social grade per ethnicity


```{r}

census$ethnicityew <- factor(census$ethnicityew, levels = seq(1,18), labels = c("English/Welsh/Scottish/Northern Irish/British", 
"Irish", 
"Gypsy or Irish Traveller", 
"Other White", 
"Mixed/multiple ethnic group: White and Black Caribbean", 
"Mixed/multiple ethnic group: White and Black African", 
"Mixed/multiple ethnic group: White and Asian", 
"Mixed/multiple ethnic group: Other Mixed", 
"Indian", 
"Pakistani", 
"Bangladeshi", 
"Chinese", 
"Other Asian", 
"African", 
"Caribbean", 
"Other Black", 
"Arab", 
"Any other ethnic group"))


census$religionew <- factor(census$religionew, levels = seq(1,9), labels = c("No religion", 
"Christian", 
"Buddhist", 
"Hindu", 
"Jewish", 
"Muslim", 
"Sikh", 
"Other religion", 
"Not stated"))

#1: white 2: Asian 3: black 4: Arab 5: other


###################### Social grade per ethnicity##################
#clean data 
cen_rel <- select(census, marstat, caseno, hlqupuk11, scgpuk11c, religionew, ethngrouped)
cen_rel$separated <- cen_rel$marstat
cen_rel <- cen_rel %>% mutate(separated = replace(separated, marstat == 2 | marstat == 3, 0))
cen_rel <- cen_rel %>% mutate(separated = replace(separated, marstat == 5 | marstat == 4, 1))
colnames(cen_rel) <- c("ID", "marstat", "education", "social_grade", "religion", "ethnicity", "separated")
#replace -9 with NA 
cen_rel <- cen_rel %>% mutate_all(funs(replace(., . == -9, NA)))
cen_rel <- cen_rel %>% filter(!is.na(social_grade), !is.na(education), !(education == 16), !is.na(religion), !(religion == "Not stated"))
dim(cen_rel)

#sort data
rel <- cen_rel %>% group_by(ethnicity, social_grade) %>% summarise(n = n(), separated = sum(separated), rel_div = (separated / n)*100) %>% mutate(sg = n/sum(n))

#plot education and social grade depending on religion
library(ggplot2)
rp <- ggplot(rel, aes(x = factor(social_grade), y = rel_div))
rp + 
  geom_col(stat = "identity", position = "dodge") +
  scale_x_discrete(labels = c("10" = "none", "11" = "O levels / GCSE", "12" = "CSE / GCSE grade 1 / A", "13" = "apprenticeship", "14" = "A levels", "15" = "university degree")) +
  scale_y_continuous(limits = c(0, 50), breaks = seq(0, 50, 3)) +
  ggtitle("Divorce rate per ethnicity and social grade") +
  xlab ("Social grade") +
  ylab ("Divore rate in %") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ ethnicity)

```

#Divorce rate per education level per ethnicity
```{r}
rel <- cen_rel %>% group_by(ethnicity, education) %>% summarise(n = n(), separated = sum(separated), rel_div = (separated / n)*100) %>% mutate(sg = n/sum(n))

#plot education and social grade depending on religion
library(ggplot2)
rp <- ggplot(rel, aes(x = factor(education), y = rel_div))
rp + 
  geom_col(stat = "identity", position = "dodge") +
  scale_x_discrete(labels = c("10" = "none", "11" = "O levels / GCSE", "12" = "CSE / GCSE grade 1 / A", "13" = "apprenticeship", "14" = "A levels", "15" = "university degree")) +
  scale_y_continuous(limits = c(0, 50), breaks = seq(0, 50, 3)) +
  ggtitle("Divorce rate per ethnicity and education level") +
  xlab ("Education level") +
  ylab ("Divore rate in %") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ ethnicity)


#1: white 2: Asian 3: black 4: Arab 5: other
```


#Education per religion

```{r}
#sort data
rel <- cen_rel %>% group_by(religion, education) %>% summarise(n = n(), separated = sum(separated), rel_div = (separated / n)*100) %>% mutate(sg = n/sum(n))

#plot education and social grade depending on religion
library(ggplot2)
rp <- ggplot(rel, aes(x = factor(education), y = rel_div))
rp + 
  geom_col(stat = "identity", position = "dodge") +
  scale_x_discrete(labels = c("10" = "none", "11" = "O levels / GCSE", "12" = "CSE / GCSE grade 1 / A", "13" = "apprenticeship", "14" = "A levels", "15" = "university degree")) +
  scale_y_continuous(limits = c(0, 50), breaks = seq(0, 50, 3)) +
  ggtitle("Divorce rate per religion and education") +
  xlab ("Education level") +
  ylab ("Divore rate in %") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ religion)
```


```{r}

rp <- ggplot(rel, aes(x = factor(education), y = rel_div))
rp + 
  geom_col(stat = "identity", position = "dodge") +
  scale_x_discrete(labels = c("10" = "none", "11" = "O levels / GCSE", "12" = "CSE / GCSE grade 1 / A", "13" = "apprenticeship", "14" = "A levels", "15" = "university degree")) +
  scale_y_continuous(limits = c(0, 50), breaks = seq(0, 50, 3)) +
  ggtitle("Divorce rate per education level") +
  xlab ("Education level") +
  ylab ("Divore rate in %") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ religion)

```

in the different religions the pattern of the divorce rate per education level differe a lot from the overall pattern. Where there is on overall a high divorce rate for people with no education, for the  muslims it is way more constant, and for Buddhists it follow a totaly different pattern. 

#divorce rate per social grade and religion

```{r}
rel <- cen_rel %>% group_by(social_grade, religion) %>% summarise(n = n(), separated = sum(separated), rel_div = (separated / n)*100) 

rp <- ggplot(rel, aes(x = factor(social_grade), y = rel_div))
rp + 
  geom_col(stat = "identity", position = "dodge") +
  scale_y_continuous(limits = c(0, 50), breaks = seq(0, 50, 3)) +
  ggtitle("Divorce rate per social grade") +
  xlab ("Social grade") +
  ylab ("Divore rate in %") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ religion)
```

```{r}
reli <- cen_rel %>% group_by(religion) %>% summarise(n = n(), separated = sum(separated), rel_div = (separated / n)*100) 


relp <- ggplot(reli, aes(x = religion, y = rel_div))
relp  +
  geom_col(fill  = "cadetblue3") +
  xlab("Religion") +
  ylab("Divorce rate in percent") +
  scale_y_continuous(limits = c(0, 40), breaks = seq(0, 40, 5)) +
  ggtitle("Divorce rate per religion")

```

#education per religion (too small sample for less represented religions)

```{r}
#clean data 
cen_rel <- select(census, caseno, marstat, hlqupuk11, scgpuk11c, religionew, ethngrouped)
cen_rel$separated <- cen_rel$marstat
cen_rel <- cen_rel %>% mutate(separated = replace(separated, marstat == 2 | marstat == 3, 0))
cen_rel <- cen_rel %>% mutate(separated = replace(separated, marstat == 5 | marstat == 4, 1))
colnames(cen_rel) <- c("ID", "marstat", "education", "social_grade", "religion", "ethnicity", "separated")
#replace -9 with NA 
cen_rel <- cen_rel %>% mutate_all(funs(replace(., . == -9, NA)))
cen_rel <- cen_rel %>% filter(!is.na(social_grade), !is.na(education), !(education == 16), !is.na(religion), !(religion == "Not stated"), !is.na(ethnicity))
dim(cen_rel)

#sort data
rel <- cen_rel %>% group_by(religion, education, social_grade) %>% summarise(n = n(), separated = sum(separated), rel_div = (separated / n)*100) %>% mutate(sg = n/sum(n))

#plot education and social grade depending on religion
library(ggplot2)
rp <- ggplot(rel, aes(x = factor(education), y = rel_div, fill = factor(social_grade)))
rp + 
  geom_col(stat = "identity", position = "dodge") +
  scale_x_discrete(labels = c("10" = "none", "11" = "O levels / GCSE", "12" = "CSE / GCSE grade 1 / A", "13" = "apprenticeship", "14" = "A levels", "15" = "university degree")) +
  scale_y_continuous(limits = c(0, 50), breaks = seq(0, 50, 3)) +
  ggtitle("Divorce rate per education level") +
  xlab ("Education level") +
  ylab ("Divore rate in %") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ religion)
```


#social grade and education
#divorce rate per social grade and education
```{r}
region <- cen_r %>% group_by(education, social_grade) %>% summarise(n = n(), separated = sum(separated), rel_div = (separated / n)*100) %>% mutate(sg = n/sum(n))


rp <- ggplot(region, aes(x = factor(social_grade), y = rel_div))
rp + 
  geom_col(stat = "identity", position = "dodge") +
  scale_x_discrete(labels = c("10" = "none", "11" = "O levels / GCSE", "12" = "CSE / GCSE grade 1 / A", "13" = "apprenticeship", "14" = "A levels", "15" = "university degree")) +
  scale_y_continuous(limits = c(0, 45), breaks = seq(0, 45, 3)) +
  ggtitle("Divorce rate per education level and region") +
  xlab ("Education level") +
  ylab ("Divore rate in %") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ education)
```


#average white person christ 
```{r}
cen_rel <- select(census, marstat, caseno, hlqupuk11, scgpuk11c, religionew, ethngrouped)
cen_rel$separated <- cen_rel$marstat
cen_rel <- cen_rel %>% mutate(separated = replace(separated, marstat == 2 | marstat == 3, 0))
cen_rel <- cen_rel %>% mutate(separated = replace(separated, marstat == 5 | marstat == 4, 1))
colnames(cen_rel) <- c("ID", "marstat", "education", "social_grade", "religion", "ethnicity", "separated")
#replace -9 with NA 
cen_rel <- cen_rel %>% mutate_all(funs(replace(., . == -9, NA)))
cen_rel <- cen_rel %>% filter(!is.na(social_grade), !is.na(education), !(education == 16), !is.na(religion), !(religion == "Not stated"))
dim(cen_rel)

#sort data
white <- cen_rel %>% filter(religion == "Christian", ethnicity ==1, social_grade == 4, education == 10) %>% summarise(n = n(), separated = sum(separated), rel_div = (separated / n)*100)

```


