---
title: "Divorce Rate in 2011 United Kingdom"
author: "Team M"
date: "October 11, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Marriage is such a big topic in human life and always a very strong predictor of people's life satisfaction and happiness level. However, not all marriages are eternal happy endings and a lot of them nowadays end in divorce. If we can understand the factors that influence the quality of marriage, we may be able to discover more preventive measures to help more people maintain a lifelong happy marriage.

To answer this question, our group examined the 2011 UK Census data and limited our scope to the marital status of a sample of British people. 

through descriptive and inferential statistics

To clarify, we define divorce rate here as (total number of divorce or separated)/(total number of marriage). Thus, we exclude all people who have not yet married, and the people who end their marriages out of other reasons such as the other partner's death.  



## Description
The dataset we examined is the 5% sample data of the 2011 UK Census. It contains 2,848,155 observations and 121 variables. Each observation represents a UK citizen
Since the dataset is very large and there are too many variables, we decided to exclusively focus on the relationship between marital status (marstat) and other variables. In brief, we would like to examine the factors that may influence people's marriage and lead to higher divorce rate. Therefore, we selected a total of 17 variables that may be highly correlated with marital status.


We can attach a table of all variables we examined here:


# Methods 
a really brief description of the R packages we used. We can load all our libraries in this page:
```{r result = "hide", message=FALSE}
library(dplyr)
library(ggplot2)
library(stargazer)
library(readr)
library(RCurl)
library(foreign)


# Load data
url <- "https://raw.githubusercontent.com/dlouhasha/TheGreatWork/master/data/filtered_dataset.csv"
filtered_data <- getURL(url)
filtered_data <- read.csv(textConnection(filtered_data))

fd <- filtered_data
fd <- as.data.frame(fd)

```


```{r, echo = FALSE, eval=FALSE}
# changing the factor names 
fd <- filtered_data
fd <- as.data.frame(fd)
fd[, 'region'] <- factor(fd[,'region'], levels = c(1, 2, 3, 4, 5, 6, 7,
                                                       8, 9, 10, 11), 
                           labels = c('North East', 'North West', 'Yorkshire and the Humber', 
                                      'East Midlands', 'West Midlands', 'East of England', 
                                      'Inner London',
                                      'Outer London', 'South East', 'South West', 'Wales'))

fd <- fd %>% mutate('binary_region_num' = ifelse(region == 'Inner London', 1, 0)) %>% mutate( 'binary_region_factor' = ifelse(region == 'Inner London', "Inner London", "Outside London"))


fd[,'marstat'] <- factor(fd[,'marstat'], level = c(1,2,3,4,5,6), 
                           labels = c('Single', 'Married', 'Same-sex', 'Separated', 
                                      'Divorced', 'Widowed'))

fd <- mutate(fd, 'binary_marstat' = ifelse(marstat == "Separated" | marstat == 'Divorced', 1, 0))

fd[,'transport']<- factor(fd[,'transport'], levels = c(1, 2, 3, 4, 5, 6, 7,
                                                           8, 9, 10, 11), 
                            labels = c('Work from home', 'metro', 'Train', 'Bus',
                                       'Taxi', 'Motocycle', 'drive a car', 'passenger in a car', 
                                       'Bicycle', 'On foot', 'Other'))

fd[,'wpzhome'] <- factor(fd[,'wpzhome'], level = c(1,2,3,4,5), 
                           labels = c('Work from home', 'Same zone', 'Outside zone, but same LA/UA district',
                                      'Outside LA/UA/LGD', 'Workplace outside UK'))

fd[,'aggdtwpew11g'] <- factor(fd[,'aggdtwpew11g'], level = c(1, 2, 3, 4, 5, 6, 7,
                                                                 8, 9, 10, 11, 12), 
                                labels = c('less than 2km', '2-5', '5-10', '10-20', 
                                           '20-40', '40-60', '60 or more', 'at home', 
                                           'No fixed place', 'Outside England and Wales but in UK', 
                                           'Outside UK', 'Offshore but within UK'))

fd <- mutate(fd, Highest_qualification = hlqupuk11)
fd[,'Highest_qualification'] <- factor(fd[,'Highest_qualification'], 
                                            levels = c(10, 11, 12, 13, 14, 15, 16), labels = c('No qualification', 'Level 1', 'Level 2', 'Apprenticeship','Level 3', 'Level 4', 'Others'))

fd[,'hours']<- factor(fd[,'hours'], level= c(1,2,3,4), labels = c('15 or less', '16-30', '31-48','49 or more'))

fd[, 'tenure'] <- factor(fd[,'tenure'], levels = c(1, 2, 3, 4, 5), 
                           labels = c('Owns outright', 'Owns with a mortgage', 'Shared ownership', 'Rents', 'Rent-free'))

fd <- fd%>% mutate(no_of_children = dpcfamuk11)%>% mutate(children_class = ifelse(dpcfamuk11==1, 'No children', ifelse(dpcfamuk11==2|dpcfamuk11==3|dpcfamuk11==4|dpcfamuk11==5|dpcfamuk11==6|dpcfamuk11==7, 'One child', ifelse(dpcfamuk11==8|dpcfamuk11==9|dpcfamuk11==10|dpcfamuk11==11|dpcfamuk11==12|dpcfamuk11==13, 'Two children', ifelse(dpcfamuk11==14|dpcfamuk11==15|dpcfamuk11==16|dpcfamuk11==17|dpcfamuk11==18|dpcfamuk11==19, 'Three or more children', NA)))))


```

# Descriptive Statistics and Correlation Tests
Simple descriptive statistis including bar charts, pie charts, density graph... etc. 
We can show a few correlation tests, chi-squre tests and graphs

## Social and Economic Status & Marriage 



## Religion and Culture & Marriage



## Age & Marriage



## Region/Geography & Marriage
An article on divorce rate by region piqued our interests (http://www.dailymail.co.uk/news/article-3201497/Wish-weren-t-ten-divorce-hot-spots-Britain-sea-Blackpool-worst-place-live-want-happy-marriage.html). The article lists the top 10 divorce hot spots and they are all coastal cities. In this paper we are keen to explore whether our data attests to the findings of that article. 

The data shows a significant peak in inner London which shows that the percentage of people in bad marriage (either divorced or separated) is much higher than elsewhere. We would like to conduct a hypothesis testing to see if region and marital status are independent. The test we adopt here is chi-squared test at 5% significance level because we are exploring the correlations between two categorical variables. 

The p-value is less than 5% and it means that the we can reject the null hypothesis that region and marital status are independent. From the residuals of chi-squared test, we see that married couples are greatly underrepresented in Inner London, but divorced and separated couples are over represented. Interestingly Inner London has a very high proportion of same-sex couples. 

We use the data to plot a mosaic plot which essentially represents the residual table graphically. 

```{r}
# select the data relevant for region
region1 <- dplyr:: select(fd, binary_marstat, age, aggdtwpew11g, region, marstat, transport, wpzhome, Highest_qualification, binary_region_factor, binary_region_num,hours, tenure)

# plot divorce rate for region
region1%>% group_by(region) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = region, y = divorce_rate))+ geom_col() + theme(axis.text.x = element_text(angle = 30, hjust = 1))

# chi-squared test
region_test <- region_count[,-1]
rownames(region_test) <- as.data.frame(region_count)[,1]
chi_region <- chisq.test(as.data.frame(region_test))
chi_region
round(chi_region$residuals, 3)

#Mosaicplot
mosaicplot(region_test, shade = TRUE, las=2,
           main = "marital status")
```


We further perform t-test to compare the divorce rate of Inner London and all the other regions. Inner London has a divorce rate of approximately 25%, higher than the average of 20% across all regions. The difference is statistically significant. 

```{r}
#mean_London <- fdata$binary_marstat[fdata$region == "Inner London"]
outside_london <- fd$binary_marstat[fd$region != "Inner London"]
t.test(fd$binary_marstat, fd$binary_marstat[fd$region == "Inner London"])
t.test(outside_london, fd$binary_marstat[fd$region == "North East"])
t.test(outside_london, fd$binary_marstat[fd$region == "North West"])
t.test(outside_london, fd$binary_marstat[fd$region == "Yorkshire and the Humber"])
t.test(outside_london, fd$binary_marstat[fd$region == "East Midlands"])
t.test(outside_london, fd$binary_marstat[fd$region == "West Midlands"])
t.test(outside_london, fd$binary_marstat[fd$region == "East of England"])
t.test(outside_london, fd$binary_marstat[fd$region == "Outer London"])
t.test(outside_london, fd$binary_marstat[fd$region == "South East"])
t.test(outside_london, fd$binary_marstat[fd$region == "South West"])
t.test(outside_london, fd$binary_marstat[fd$region == "Wales"])

```

The article aforementioned states that the factors contributing to the high divorce rate in coastal regions are due to the high deprivation and possibly sheer boredoem in winter. Our data shows otherwise. London is anything but a deprived and boring city. What could have contributed to the different in findings? 

Secondly, we suspect there are other more salient influencing factors and this secton of our paper will explore that. We explore the variables by looking at how the divorce rates for these variables vary across region. We explored education, social grade, age, hours worked, number of children in family, tenure of dwelling, ethinicity, religion and transport. Other than age, the pattern is very similar across regions. We observe a significantly high divorce rate for people aged 50 and above in Inner London. However Inner London is a very young city with mean age of residents at around 47, and median age 45 (population mean and median are both 52). 

We further break down the population into two groups: age below 50 and age 50 and above. In the first group we no longer see that any significant difference in divorce rates across regions, but the difference becomes more apparent in the second group. The observation suggests a larger proportion of people aged 50 and above are divorced or separated in Inner London than elsewhere. 

According to this research by Susan L. Brown and I-Fen Lin (18 Oct 2012).  https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3478728/
Various factors contribute to the increasing divorce rate among older people. A better economic security, and more social interactions with people could have contributed to that as well. Or it is also possible that many people are in their second marriage in their 50s and 60s and high-order marriages are more likely to end up in divorces than first orders. Or it could also be possible that older people in Inner London are less likely to remarry after they are divorced, just like London has the highest percentage of singles aged 30 and above. 


18 July 2017  By ONS Digital
Office for National Statistics reported that while the number of divorces drop by 28% from 2005 to 2015, the number of men divorcing aged above 65 went up by 23% and 38% for women.  (https://visual.ons.gov.uk/marriage-and-divorce-on-the-rise-at-65-and-over/)




```{r region_vs_others, echo = FALSE, eval= FALSE}

fd %>% group_by(region, age) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = age, y = divorce_rate))+ geom_col() + facet_wrap(~region)

fd %>% filter(!is.na(children_class))%>% group_by(region, children_class) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = children_class, y = divorce_rate))+ geom_col() + facet_wrap(~region)+ theme(axis.text.x = element_text(angle = 60, hjust = 1))  

fd %>% filter(!is.na(Highest_qualification) & Highest_qualification != 'Others') %>%group_by(region, Highest_qualification) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = Highest_qualification, y = divorce_rate))+ geom_col() + facet_wrap(~region)+ theme(axis.text.x = element_text(angle = 60, hjust = 1)) 


fd %>% filter(!is.na(hours)) %>%group_by(region, hours) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = hours, y = divorce_rate))+ geom_col() + facet_wrap(~region)+ theme(axis.text.x = element_text(angle = 60, hjust = 1))  

fd %>% filter(!is.na(ethnicityew)) %>%group_by(region, ethnicityew) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = ethnicityew, y = divorce_rate))+ geom_col() + facet_wrap(~region)+ theme(axis.text.x = element_text(angle = 60, hjust = 1))  

fd %>% filter(!is.na(religionew)) %>%group_by(region, religionew) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = religionew, y = divorce_rate))+ geom_col() + facet_wrap(~region)+ theme(axis.text.x = element_text(angle = 60, hjust = 1))  


fd %>% filter(!is.na(tenure)) %>%group_by(region, tenure) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = tenure, y = divorce_rate))+ geom_col() + facet_wrap(~region)+ theme(axis.text.x = element_text(angle = 60, hjust = 1))  


fd %>% filter(!is.na(wpzhome)) %>%group_by(region, wpzhome) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = wpzhome, y = divorce_rate))+ geom_col() + facet_wrap(~region)+ theme(axis.text.x = element_text(angle = 60, hjust = 1))  

```

```{r region_vs_age}

fd %>% group_by(region, age) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = age, y = divorce_rate))+ geom_col() + facet_wrap(~region)

fd %>% filter(!is.na(age))%>% ggplot(aes(x = age))+ geom_histogram() + facet_wrap(~region)

region1%>% filter(age <50) %>% group_by(region) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = region, y = divorce_rate))+ geom_col() + theme(axis.text.x = element_text(angle = 30, hjust = 1))


region1%>% filter(age >=50) %>% group_by(region) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = region, y = divorce_rate))+ geom_col() + theme(axis.text.x = element_text(angle = 30, hjust = 1))

```

### Dependent children and tenure

```{r}
family_compo <- fd%>% select(dpcfamuk11, tenure, binary_marstat, binary_region_factor) %>% mutate(no_of_children = dpcfamuk11)%>% mutate(children_class = ifelse(dpcfamuk11==1, 'No children', ifelse(dpcfamuk11==2|dpcfamuk11==3|dpcfamuk11==4|dpcfamuk11==5|dpcfamuk11==6|dpcfamuk11==7, 'One child', ifelse(dpcfamuk11==8|dpcfamuk11==9|dpcfamuk11==10|dpcfamuk11==11|dpcfamuk11==12|dpcfamuk11==13, 'Two children', ifelse(dpcfamuk11==14|dpcfamuk11==15|dpcfamuk11==16|dpcfamuk11==17|dpcfamuk11==18|dpcfamuk11==19, 'Three or more children', NA)))))


family_compo %>% select(binary_marstat, tenure, binary_region_factor, children_class) %>% filter(!is.na(children_class))%>% group_by(children_class)%>% summarise(divorce_rate = mean(binary_marstat))%>%ggplot(aes(x = children_class, y = divorce_rate))+ geom_col() + theme(axis.text.x = element_text(angle = 60, hjust = 1))  

```


## Dependent Children and Social Grade

```{r}
family_compo <- fd%>% select(dpcfamuk11, tenure, scgpuk11c, binary_marstat, binary_region_factor) %>% mutate(no_of_children = dpcfamuk11)%>% mutate(children_class = ifelse(dpcfamuk11==1, 'No children', ifelse(dpcfamuk11==2|dpcfamuk11==3|dpcfamuk11==4|dpcfamuk11==5|dpcfamuk11==6|dpcfamuk11==7, 'One child', ifelse(dpcfamuk11==8|dpcfamuk11==9|dpcfamuk11==10|dpcfamuk11==11|dpcfamuk11==12|dpcfamuk11==13, 'Two children', ifelse(dpcfamuk11==14|dpcfamuk11==15|dpcfamuk11==16|dpcfamuk11==17|dpcfamuk11==18|dpcfamuk11==19, 'Three or more children', NA)))))

family_compo %>% select(binary_marstat, tenure, scgpuk11c, binary_region_factor, children_class) %>% filter(scgpuk11c != -9) %>% filter(!is.na(children_class)) %>% group_by(scgpuk11c, children_class)%>% summarise(divorce_rate = mean(binary_marstat))%>%ggplot(aes(x = scgpuk11c, y = divorce_rate, fill = children_class))+ geom_col(position = 'dodge') + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + labs(x = 'Social Class')
```





## Health & Marriage
```{r message = FALSE, warning = FALSE, echo = FALSE}
#This should be made universal and not distinct to any one local machine

url <- "https://raw.githubusercontent.com/dlouhasha/TheGreatWork/master/data/filtered_dataset.csv"
filtered_data <- getURL(url)
filtered_data <- read.csv(textConnection(filtered_data))

fd <- filtered_data
mine <- select(fd, X1, disability, marstat, binary_marstat)

mine$disability <- factor(mine$disability, levels = c(1,2,3), labels = c("Limit a lot", "Limit a little", "not limited") )
mine$binary_marstat <- factor(mine$binary_marstat, levels = c(0,1), labels = c("married", "divorced"))
chisq.test(mine$binary_marstat,mine$disability)

```

A chi-square test of the two variables indicate that the correlation between marital status and long-term health problem is significant. This finding suggests that the two variables are not independent of each other. Long-term health problem will influence marital status. 

```{r echo = FALSE}
mine2 <- select(fd, disability, marstat, binary_marstat)
mine2$disability <- factor(mine2$disability, levels = c(1,2,3), labels = c("Limit a lot", "Limit a little", "not limited") )
mine2 <- mine2 %>% group_by(disability) %>% summarise(divorce_rate = mean(binary_marstat))
ggplot(mine, aes(disability, fill =binary_marstat))+geom_bar(position = "fill")
ggplot(mine2, aes(disability, divorce_rate))+geom_bar(stat = "identity")
```
The graph shows that the less severe the long-term health problem, the less the divorced rate.



# Logistic Regression
A comprehensive regression analysis of all correlated variables



# Hypothesis Testing and Prediction Model (Helene's)



# Conclusion

Our findings will go here

```{r Findings}

```

# References

Main Github Repository: https://github.com/dlouhasha/TheGreatWork

# Appendix
A list of all variables and descriptions