---
title: "Divorce Rate in 2011 United Kingdom"
author: "Team M"
date: "October 11, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

Marriage is such a big topic in human life and always a very strong predictor of people's life satisfaction and happiness level. However, not all marriages are eternal happy endings and a lot of them nowadays end in divorce. If we can understand the factors that influence the quality of marriage, we may be able to discover more preventive measures to help more people maintain a lifelong happy marriage.

To answer this question, our group examined the 2011 UK Census data and limited our scope to the marital status of a sample of British people. 

through descriptive and inferential statistics

To clarify, we define divorce rate here as (total number of divorce or separated)/(total number of marriage). Thus, we exclude all people who have not yet married, and the people who end their marriages out of other reasons such as the other partner's death.  



## Description
The dataset we examined is the 5% sample data of the 2011 UK Census. It contains 2,848,155 observations and 121 variables. Each observation represents a UK citizen
Since the dataset is very large and there are too many variables, we decided to exclusively focus on the relationship between marital status (marstat) and other variables. In brief, we would like to examine the factors that may influence people's marriage and lead to higher divorce rate. Therefore, we selected a total of 17 variables that may be highly correlated with marital status.


We can attach a table of all variables we examined here:


# Methods 
a really brief description of the R packages we used. We can load all our libraries in this page:
```{r result = "hide", message=FALSE}
library(dplyr)
library(ggplot2)
library(stargazer)
library(readr)
library(RCurl)
library(foreign)


# Load data
url <- "https://raw.githubusercontent.com/dlouhasha/TheGreatWork/master/data/filtered_dataset.csv"
filtered_data <- getURL(url)
filtered_data <- read.csv(textConnection(filtered_data))

fd <- filtered_data
fd <- as.data.frame(fd)

```


```{r, echo = FALSE}
# changing the factor names 
fd[, 'region'] <- factor(fd[,'region'], levels = c(1, 2, 3, 4, 5, 6, 7,
                                                       8, 9, 10, 11), 
                           labels = c('North East', 'North West', 'Yorkshire and the Humber', 
                                      'East Midlands', 'West Midlands', 'East of England', 
                                      'Inner London',
                                      'Outer London', 'South East', 'South West', 'Wales'))

fd <- fd %>% mutate('binary_region_num' = ifelse(region == 'Inner London', 1, 0)) %>% mutate( 'binary_region_factor' = ifelse(region == 'Inner London', "Inner London", "Outside London"))


fd[,'marstat'] <- factor(fd[,'marstat'], level = c(1,2,3,4,5,6), 
                           labels = c('Single', 'Married', 'Same-sex', 'Separated', 
                                      'Divorced', 'Widowed'))

fd <- mutate(fd, 'binary_marstat' = ifelse(marstat == "Separated" | marstat == 'Divorced', 1, 0))

fd[,'transport']<- factor(fd[,'transport'], levels = c(1, 2, 3, 4, 5, 6, 7,
                                                           8, 9, 10, 11), 
                            labels = c('Work from home', 'metro', 'Train', 'Bus',
                                       'Taxi', 'Motocycle', 'drive a car', 'passenger in a car', 
                                       'Bicycle', 'On foot', 'Other'))

fd[,'wpzhome'] <- factor(fd[,'wpzhome'], level = c(1,2,3,4,5), 
                           labels = c('Work from home', 'Same zone', 'Outside zone, but same LA/UA district',
                                      'Outside LA/UA/LGD', 'Workplace outside UK'))

fd[,'aggdtwpew11g'] <- factor(fd[,'aggdtwpew11g'], level = c(1, 2, 3, 4, 5, 6, 7,
                                                                 8, 9, 10, 11, 12), 
                                labels = c('less than 2km', '2-5', '5-10', '10-20', 
                                           '20-40', '40-60', '60 or more', 'at home', 
                                           'No fixed place', 'Outside England and Wales but in UK', 
                                           'Outside UK', 'Offshore but within UK'))

fd <- mutate(fd, Highest_qualification = hlqupuk11)
fd[,'Highest_qualification'] <- factor(fd[,'Highest_qualification'], 
                                            levels = c(10, 11, 12, 13, 14, 15, 16), labels = c('No qualification', 'Level 1', 'Level 2', 'Apprenticeship','Level 3', 'Level 4', 'Others'))

fd[,'hours']<- factor(fd[,'hours'], level= c(1,2,3,4), labels = c('15 or less', '16-30', '31-48','49 or more'))

fd[, 'tenure'] <- factor(fd[,'tenure'], levels = c(1, 2, 3, 4, 5), 
                           labels = c('Owns outright', 'Owns with a mortgage', 'Shared ownership', 'Rents', 'Rent-free'))

fd <- fd%>% mutate(no_of_children = dpcfamuk11)%>% mutate(children_class = ifelse(dpcfamuk11==1, 'No children', ifelse(dpcfamuk11==2|dpcfamuk11==3|dpcfamuk11==4|dpcfamuk11==5|dpcfamuk11==6|dpcfamuk11==7, 'One child', ifelse(dpcfamuk11==8|dpcfamuk11==9|dpcfamuk11==10|dpcfamuk11==11|dpcfamuk11==12|dpcfamuk11==13, 'Two children', ifelse(dpcfamuk11==14|dpcfamuk11==15|dpcfamuk11==16|dpcfamuk11==17|dpcfamuk11==18|dpcfamuk11==19, 'Three or more children', NA)))))

fd <- fd %>% mutate(youngest_child_age = ifelse(dpcfamuk11==1, 'No children', ifelse(dpcfamuk11==2|dpcfamuk11==8|dpcfamuk11==14, 'youngest aged 0-4', ifelse(dpcfamuk11==3|dpcfamuk11==9|dpcfamuk11==15, 'youngest aged 5-7', ifelse(dpcfamuk11==4|dpcfamuk11==10|dpcfamuk11==16, 'youngest aged 8-9', ifelse(dpcfamuk11==5|dpcfamuk11==11|dpcfamuk11==17, 'youngest aged 10-11', ifelse(dpcfamuk11==6|dpcfamuk11==12|dpcfamuk11==18, 'youngest aged 12-15', ifelse(dpcfamuk11==7|dpcfamuk11==13|dpcfamuk11==19, 'youngest aged 16-18', NA))))))))

```

# Descriptive Statistics and Correlation Tests
Simple descriptive statistis including bar charts, pie charts, density graph... etc. 
We can show a few correlation tests, chi-squre tests and graphs

## Social and Economic Status & Marriage 



## Religion and Culture & Marriage



## Age & Marriage



## Region/Geography & Marriage
An article on divorce rate by region piqued our interests (http://www.dailymail.co.uk/news/article-3201497/Wish-weren-t-ten-divorce-hot-spots-Britain-sea-Blackpool-worst-place-live-want-happy-marriage.html). It lists the top 10 divorce hot spots and they are all coastal cities. The article states that the factors contributing to the high divorce rate in coastal regions are due to the high deprivation and possibly sheer boredoem in winter. In this paper we are keen to explore whether our data attests to the findings of that article. 

Our data shows a significant peak in inner London which shows that the percentage of people in bad marriage (either divorced or separated) is much higher than elsewhere. This is the opposite to the aforementioned article. London is anything but a deprived and boring city. We would like to conduct a hypothesis testing to see (1) if region and marital status are independent, (2) if London has the highest divorce rate and (3) if so what the influencing factors are. The test we adopt here is chi-squared test to explore the correlations between region and marital status and t-test for testing of difference in divorce rates across regions. 

The p-values are both less than 5% and it means that the we can reject the null hypothesis that region and marital status are independent. From the residuals of chi-squared test, we see that married couples are greatly underrepresented in Inner London, but divorced and separated couples are over represented. 

```{r, echo = FALSE}
# select the data relevant for region
region1 <- dplyr:: select(fd, binary_marstat, age, aggdtwpew11g, region, marstat, transport, wpzhome, Highest_qualification, binary_region_factor, binary_region_num,hours, tenure)

region_count<- region1 %>% select(marstat, region) %>% group_by(region, marstat)%>% summarise(count = n())%>% tidyr::spread(key = marstat, value = count) %>% mutate(divorce_rate = (Divorced + Separated)/(Divorced + Married + Separated))

```

```{r region_plot, fig.width=3,fig.height=3}
# plot divorce rate for region
region1%>% group_by(region) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = region, y = divorce_rate))+ geom_col() + theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

```{r}
# chi-squared test
region_test <- region_count[,-1]
rownames(region_test) <- as.data.frame(region_count)[,1]
chi_region <- chisq.test(as.data.frame(region_test))
chi_region

```

Chi-squared test residuals on regional marital status analysis
```{r}
round(chi_region$residuals, 3)['Inner London',]
```

We further perform t-test to compare the divorce rate of Inner London and all the other regions. Inner London has a divorce rate of approximately 25%, higher than the average of 20% across all regions. The difference is statistically significant. 

```{r region_ttest, echo = FALSE}
t.test(fd$binary_marstat, fd$binary_marstat[fd$region == "Inner London"])
```

```{r, echo = FALSE, eval = FALSE}
#mean_London <- fdata$binary_marstat[fdata$region == "Inner London"]
outside_london <- fd$binary_marstat[fd$region != "Inner London"]

t.test(outside_london, fd$binary_marstat[fd$region == "North East"])
t.test(outside_london, fd$binary_marstat[fd$region == "North West"])
t.test(outside_london, fd$binary_marstat[fd$region == "Yorkshire and the Humber"])
t.test(outside_london, fd$binary_marstat[fd$region == "East Midlands"])
t.test(outside_london, fd$binary_marstat[fd$region == "West Midlands"])
t.test(outside_london, fd$binary_marstat[fd$region == "East of England"])
t.test(outside_london, fd$binary_marstat[fd$region == "Outer London"])
t.test(outside_london, fd$binary_marstat[fd$region == "South East"])
t.test(outside_london, fd$binary_marstat[fd$region == "South West"])
t.test(outside_london, fd$binary_marstat[fd$region == "Wales"])

```

Could the place people choose to live affect their marriage satisfaction, or is it merely an artifact as a result of distributions of other influencing factors? We plot the divorce rates against other variables in our list and wrap them by region. If the variables have no impact, we should observe the same pattern across all regions, with London having a slightly higher divorce rate. 

Findings: we observe a significantly high divorce rate for people aged 50 and above in Inner London, while inner London is a very young city with mean age of residents at around 47, and median age 45 (population mean and median are both 52). To explore it further we break down the population into two groups: age below 50 and age 50 and above. In the first group we no longer see any significant difference in divorce rates across regions, but the difference becomes more apparent in the second group. The observation suggests a larger proportion of people aged 50 and above are divorced or separated in Inner London than elsewhere. 

According to Office for National Statistics in their online publication in July 2017, while the number of divorces drop by 28% from 2005 to 2015, the number of men divorcing aged above 65 went up by 23% and 38% for women.  (https://visual.ons.gov.uk/marriage-and-divorce-on-the-rise-at-65-and-over/). 

Various factors contribute to the increasing divorce rate among older people (Susan L. Brown and I-Fen Lin 18 Oct 2012, https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3478728/). A better economic security, and more social interactions with people are among them. More significanty observed in their data is that people in higher order marriages are more likely to end up in divorces than people in first marriage (53.46% compared to 46.54%). And people in their 50s and 60s are also more likely to be in higher order marriages compared to younger people. 

However we must be careful not to confuse the status of being divorced or separated with the instance of getting a divorce. In our data we look at the percentage of people who are divorced or separated at the time of the survey, without information on when and where the divorce occured. It could also be possible that people in inner London are less likely to remarry after they are divorced, just like London has the highest percentage of singles aged 30 and above. London is fun and people could be just having a good time being single or in relationships, without rushing into any marriage. Above all, London is the number one Tinder hotspots. 


```{r region_vs_others, echo = FALSE, eval= FALSE}
# divorce rate by variables and wrapped by region

#divorce rate by age 
fd %>% group_by(region, age) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = age, y = divorce_rate))+ geom_col() + facet_wrap(~region)

#divorce rate by children_class
fd %>% filter(!is.na(children_class))%>% group_by(region, children_class) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = children_class, y = divorce_rate))+ geom_col() + facet_wrap(~region)+ theme(axis.text.x = element_text(angle = 60, hjust = 1))  

#divorce rate by Highest_qualification
fd %>% filter(!is.na(Highest_qualification) & Highest_qualification != 'Others') %>%group_by(region, Highest_qualification) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = Highest_qualification, y = divorce_rate))+ geom_col() + facet_wrap(~region)+ theme(axis.text.x = element_text(angle = 60, hjust = 1)) 

#divorce rate by hours worked
fd %>% filter(!is.na(hours)) %>%group_by(region, hours) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = hours, y = divorce_rate))+ geom_col() + facet_wrap(~region)+ theme(axis.text.x = element_text(angle = 60, hjust = 1))  

#divorce rate by ethnicity
fd %>% filter(!is.na(ethnicityew)) %>%group_by(region, ethnicityew) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = ethnicityew, y = divorce_rate))+ geom_col() + facet_wrap(~region)+ theme(axis.text.x = element_text(angle = 60, hjust = 1))  

#divorce rate by religion
fd %>% filter(!is.na(religionew)) %>%group_by(region, religionew) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = religionew, y = divorce_rate))+ geom_col() + facet_wrap(~region)+ theme(axis.text.x = element_text(angle = 60, hjust = 1))  

#divorce rate by tenure
fd %>% filter(!is.na(tenure)) %>%group_by(region, tenure) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = tenure, y = divorce_rate))+ geom_col() + facet_wrap(~region)+ theme(axis.text.x = element_text(angle = 60, hjust = 1))  

#divorce rate by distance to work
fd %>% filter(!is.na(wpzhome)) %>%group_by(region, wpzhome) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = wpzhome, y = divorce_rate))+ geom_col() + facet_wrap(~region)+ theme(axis.text.x = element_text(angle = 60, hjust = 1))  

```

```{r region_vs_age}
# divorce rate by age and wrapped by regions
fd %>% group_by(region, age) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = age, y = divorce_rate))+ geom_col() + facet_wrap(~region)
```

```{r, echo = FALSE, eval = FALSE}
#distribution of age wrapped by region
fd %>% filter(!is.na(age))%>% ggplot(aes(x = age))+ geom_histogram() + facet_wrap(~region)

#divorce rate by region for aged below 50
region1%>% filter(age <50) %>% group_by(region) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = region, y = divorce_rate))+ geom_col() + theme(axis.text.x = element_text(angle = 30, hjust = 1))

#divorce rate by region for aged 50 and above
region1%>% filter(age >=50) %>% group_by(region) %>% summarise(divorce_rate = mean(binary_marstat))%>% ggplot(aes(x = region, y = divorce_rate))+ geom_col() + theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

### Dependent children

When people get married they are often in love and happy, most married couples look forward parenthood and having children could even strengthen their marital bond. However empirical study shows that the effect of children on marital stability could be very complex. Children increase the stability of marriage only when they are preschool-aged (Becker et al. 1977), as they grow older, their binding force in a marriage weakens; the greater number of children, the smaller possibility of divorce, but the relationship is inversely quadratic rather than linear suggesting the stabilizing effect is the strongest with the first and second child (Xu, Yu and Qiu 2015). 

The birth of an infant often brings couples back together who would have otherwise divorced. Experiencing divorce often place young children at risk for a variety of problems, which may persist into adulthood (Morrisox and Coiro, 1999). Married couples may choose to stay in marriges for the wellbeing of their children but when children are old enough they decide to part their ways. 

We are keen to explore if we observe the same pattern in our data (1) as the age of youngest child grows, the more likely the couple will get divorced/separated (2) the more children a couple have, the less likely they are to be divorced/separated. 


Becker, Gary S., Elizabeth Landes, and Robert T. Michael. 1977. "An Economic Analysis of Marital Instability." Journal of Political Economy 85:1141-87.

Waite, Linda J, and Lee A Lillard. 1991. Children and marital disruption. American Journal of Sociology 96(4):930–953. 

Qi Xu1*, Jianning Yu2 and Zeqi Qiu3. 2015, The impact of children on divorce risk. Xu et al. The Journal of Chinese Sociology (2015) 2:1

Donna Ruane Morrisox and Mary Jo Coiro, Parental Conflict and Marital Disruption: Do Children Benefit When High-Conflict Marriages Are Dissolved?. Journal of Marriage and the Family 6 1 (August 1999): 626-637


```{r}
family_compo <- fd%>% select(youngest_child_age, dpcfamuk11, tenure, binary_marstat, binary_region_factor, scgpuk11c, children_class) 

%>% mutate(no_of_children = dpcfamuk11)%>% mutate(children_class = ifelse(dpcfamuk11==1, 'No children', ifelse(dpcfamuk11==2|dpcfamuk11==3|dpcfamuk11==4|dpcfamuk11==5|dpcfamuk11==6|dpcfamuk11==7, 'One child', ifelse(dpcfamuk11==8|dpcfamuk11==9|dpcfamuk11==10|dpcfamuk11==11|dpcfamuk11==12|dpcfamuk11==13, 'Two children', ifelse(dpcfamuk11==14|dpcfamuk11==15|dpcfamuk11==16|dpcfamuk11==17|dpcfamuk11==18|dpcfamuk11==19, 'Three or more children', NA)))))


family_compo %>% select(binary_marstat, tenure, binary_region_factor, children_class, dpcfamuk11) %>% filter(!is.na(children_class))%>% group_by(children_class)%>% summarise(divorce_rate = mean(binary_marstat))%>%ggplot(aes(x = children_class, y = divorce_rate))+ geom_col() + theme(axis.text.x = element_text(angle = 60, hjust = 1))  

family_compo %>%filter(dpcfamuk11 != -9)%>% select(binary_marstat, tenure, binary_region_factor, children_class, dpcfamuk11) %>% filter(!is.na(dpcfamuk11))%>% group_by(children_class, dpcfamuk11)%>% summarise(divorce_rate = mean(binary_marstat))%>%ggplot(aes(x = dpcfamuk11, y = divorce_rate, fill = children_class))+ geom_col() + theme(axis.text.x = element_text(angle = 60, hjust = 1))  



family_compo <- fd%>% select(dpcfamuk11, tenure, scgpuk11c, binary_marstat, binary_region_factor) %>% mutate(no_of_children = dpcfamuk11)%>% mutate(children_class = ifelse(dpcfamuk11==1, 'No children', ifelse(dpcfamuk11==2|dpcfamuk11==3|dpcfamuk11==4|dpcfamuk11==5|dpcfamuk11==6|dpcfamuk11==7, 'One child', ifelse(dpcfamuk11==8|dpcfamuk11==9|dpcfamuk11==10|dpcfamuk11==11|dpcfamuk11==12|dpcfamuk11==13, 'Two children', ifelse(dpcfamuk11==14|dpcfamuk11==15|dpcfamuk11==16|dpcfamuk11==17|dpcfamuk11==18|dpcfamuk11==19, 'Three or more children', NA)))))

family_compo %>% select(dpcfamuk11, binary_marstat, tenure, scgpuk11c, binary_region_factor, children_class) %>% filter(scgpuk11c != -9) %>% filter(!is.na(children_class)) %>% group_by(dpcfamuk11, children_class)%>% summarise(divorce_rate = mean(binary_marstat))%>%ggplot(aes(x = dpcfamuk11, y = divorce_rate, fill = children_class))+ geom_col(position = 'dodge') + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + labs(x = 'Number of children in family and the youngest child age')


```




## Health & Marriage
```{r message = FALSE, warning = FALSE, echo = FALSE}
#This should be made universal and not distinct to any one local machine
mine <- select(fd, disability, marstat, binary_marstat)

mine$disability <- factor(mine$disability, levels = c(1,2,3), labels = c("Limit a lot", "Limit a little", "not limited") )
mine$binary_marstat <- factor(mine$binary_marstat, levels = c(0,1), labels = c("married", "divorced"))
chisq.test(mine$binary_marstat,mine$disability)

```

A chi-square test of the two variables indicate that the correlation between marital status and long-term health problem is significant. This finding suggests that the two variables are not independent of each other. Long-term health problem will influence marital status. 

```{r echo = FALSE}
mine2 <- select(fd, disability, marstat, binary_marstat)
mine2$disability <- factor(mine2$disability, levels = c(1,2,3), labels = c("Limit a lot", "Limit a little", "not limited") )
mine2 <- mine2 %>% group_by(disability) %>% summarise(divorce_rate = mean(binary_marstat))
ggplot(mine, aes(disability, fill =binary_marstat))+geom_bar(position = "fill")
ggplot(mine2, aes(disability, divorce_rate))+geom_bar(stat = "identity")
```
The graph shows that the less severe the long-term health problem, the less the divorced rate.



# Logistic Regression
A comprehensive regression analysis of all correlated variables



# Hypothesis Testing and Prediction Model (Helene's)



# Conclusion

Our findings will go here

```{r Findings}

```

# References

Main Github Repository: https://github.com/dlouhasha/TheGreatWork

# Appendix
A list of all variables and descriptions
