---
title: "Copy of main report"
author: "Hadi Kadi"
date: "October 12, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Age & Marriage
Contains information about the age of the observation (person) in years. This variable ranges from 0 to 90 and is treated as a continuous variable. Ages 70, 75, 80, and 85 are clustered into groups each ranging 5 years (i.e. 70 through 74, 75 through 79, etc...). Furthermore, the 90 age group conatins ages 90 and above.

For the purposes of this analysis, and to calculate the divorce rate by age, the clustered age groups were equally distibuted among the number of ages the group contains. (e.g the 70 through 74 age group was divided into 5 seperate age groups 70, 71, 72, 73, and 74 each inherenting the parent group's divorce rate).

Exhibit 1: Density of un-modified age

The chart below, depicts the density distribution of ages in the dataset before any modifications were applied. As expected, instances of the observations' ages are relatively equal among the dataset, which should ensure a fair representation of all age groups.
```{r}
#need to add in original dataset to show density of original age distribution
ggplot(data = fd, aes(x = age)) + geom_density()
```

Exhibit 2: Divorce rate by age:

The chart above shows the divorce rate (as defined in the introduction) for each of the 83 age groups (18 to 100). 
The divorce rate for observations aged 18, 19, and 20 is significantly higher than observations in their 20's. The rationale for this observation is the scarcity (nominally) of married people who are 20 and less, which results in a large relative proportion of divorced people when compared to other age groups.
The age with the highest divorce rate in the dataset is 48. This is likely due to the fact that younger people are less likely to be divorced, as those would be early-on in their marriages. Furthermore, older people have a greater chance of being being remarried.

```{r}

data2 <-  fd %>% filter(marstat == 2 | marstat == 3 | marstat == 4 | marstat == 5) %>% filter(age >= 18) %>% mutate(seperated = marstat) %>% mutate(seperated = replace(seperated, marstat!= 5 | marstat != 4, 0)) %>% mutate(seperated = replace(seperated, marstat == 5 | marstat == 4, 1))
#new dataframe containing the summarised count by age
total_group <- data2 %>% dplyr::select(age, seperated) %>% group_by(age) %>% summarise(count = n())
total_group_full <- total_group[1:52,]
dummy1 <- data.frame(age = 70:100)
dummy1$count <- rep(1,31)
dummy1$count[1:5] <- rep(round(total_group$count[53] / 5), 5)
dummy1$count[6:10] <- rep(round(total_group$count[54] / 5), 5)
dummy1$count[11:15] <- rep(round(total_group$count[55] / 5), 5)
dummy1$count[16:20] <- rep(round(total_group$count[56] / 5), 5)
dummy1$count[21:31] <- rep(round(total_group$count[57] / 11), 11)
new_total_group <- rbind.data.frame(total_group_full, dummy1)

#create a new sub-data frame from data2 grouped by age and seperated status and summarised 
divorced <-  data2 %>% dplyr::select(age, seperated) %>% filter(seperated == 1) %>% group_by(age, seperated) %>% summarise(n = n()) %>% ungroup() %>% group_by(seperated) %>% mutate(prop = n / sum(n) * 100)

#new divorced_full dataframe that contains all age groups above 70 i.e (70 to 100) since the orignal dataset groups ges above 70 by increments of 5. Count of divorced people was divided equally among the five tranches
divorced_full <- divorced[1:52,]
dummy <- data.frame(age = 70:100)
dummy$seperated <- rep(1,31)
dummy$n <- rep(1,31)
dummy$prop <- rep(1,31)
dummy$n[1:5] <- rep(divorced$n[53] / 5, 5)
dummy$prop[1:5] <- rep(divorced$prop[53] / 5, 5)
dummy$n[6:10] <- rep(divorced$n[54] / 5, 5)
dummy$prop[6:10] <- rep(divorced$prop[54] / 5, 5)
dummy$n[11:15] <- rep(divorced$n[55] / 5, 5)
dummy$prop[11:15] <- rep(divorced$prop[55] / 5, 5)
dummy$n[16:20] <- rep(divorced$n[56] / 5, 5)
dummy$prop[16:20] <- rep(divorced$prop[56] / 5, 5)
dummy$n[21:31] <- rep(divorced$n[57] / 11, 11)
dummy$prop[21:31] <- rep(divorced$prop[57] / 11, 11)
new_divorced <- rbind.data.frame(divorced_full, dummy)

#create new data.frame to calculate the divorce rate per age. This is done by dividing the count of divorced people from divorced data frame by the respective rows from the total group data frame 
divorce_by_age <- data.frame(prop = c(1:83), age = new_divorced$age)
divorce_by_age$prop <- rep(1, 83)
divorce_by_age$prop <- new_divorced$n / new_total_group$count * 100

ggplot(data = divorce_by_age, aes(x = age, y = prop)) + geom_line()
```

Exhibit 3: Proportion of divorced people by age

The chart above shows the proportion of divorced people for each age group. The chart shows a curve resembling a normal distribution.
```{r}
ggplot(data = new_divorced, aes(x = age, y = prop)) + geom_line()
```

Exhibit 4: Divorce rate by age and sex

The chart below is an exhibit of divorce rates by age for each sex. The general trend presented is a consistintly higher divorce rate for women across all age groups. This seems to suggest that women are more likely to be divorced than men on average.
A t test of mean divorce rate between men and women reveals a statistically significant difference between the mean divorce rate of men and women.

```{r}
data3 <-  data2 %>% filter(age > 22 & age < 70)
total_group1 <- data3 %>% dplyr::select(age, seperated, sex) %>% group_by(age, sex) %>% summarise(count = n())
divorce_by_age_bysex <- data.frame(age = total_group1$age)
divorce_by_age_bysex$prop <- divorced_bysex$n / total_group1$count * 100
divorce_by_age_bysex$sex <- divorced_bysex$sex
divorce_by_age_bysex$sex <- factor(divorce_by_age_bysex$sex, levels = c(1,2), labels = c('Male', 'Female'))
ggplot(data = divorce_by_age_bysex, aes(x = age, y = prop, fill = sex)) + geom_col(position = 'dodge') + ylab("%")

male <- divorce_by_age_bysex %>% filter(sex == 'Male')
female <- divorce_by_age_bysex %>% filter(sex == 'Female')
t.test(male$prop, female$prop)


```

# Logistic Regression
A comprehensive regression analysis of all correlated variables

The variables presented thus far in this paper are the variables hypothesised to have the greatest effect on a person's likliness of being divorced in a certain society.

As the dependant variable in question is a dichotomous variable (divorced or not divorced), a logistic regression was performed where the divorced status is the dependant variable, and the variables mentioned above are the independant variables.

A logistic model is used to predict the probability of a binary response based on a set of other variables.

Regression formula:
divorced status = intercept + b1age + b2education_level + b3number_of_children + b4religion + b5london_or_not +...


As a measure of the overall model's explanatory powe, three main measures were use:
1) Akaike Information Criterion 
2) c-statistic or AUC (area under ROC curve)
3) The accuracy of the predictions above a certain threshold (the mean divorce rate for our purposes)

The best fit was determined by adding and removing the variables above until the best combination of the aforementioned metrics was obtained.

*insert model here*
