---
title: "Copy of main report"
author: "Hadi Kadi"
date: "October 12, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Age & Marriage
Contains information about the age of the observation (person) in years. This variable ranges from 0 to 90 and is treated as a continuous variable. Ages 70, 75, 80, and 85 are clustered into groups each ranging 5 years (i.e. 70 through 74, 75 through 79, etc...). Furthermore, the 90 age group conatins ages 90 and above.

For the purposes of this analysis, and to calculate the divorce rate by age, the clustered age groups were equally distibuted among the number of ages the group contains. (e.g the 70 through 74 age group was divided into 5 seperate age groups 70, 71, 72, 73, and 74 each inherenting the parent group's divorce rate).

Exhibit 1: Density of un-modified age

The chart below, depicts the density distribution of ages in the dataset before any modifications were applied. As expected, instances of the observations' ages are relatively equal among the dataset, which should ensure a fair representation of all age groups.
```{r}
#need to add in original dataset to show density of original age distribution
ggplot(data = fd, aes(x = age)) + geom_density(fill = 'grey') + ylab("Density") + xlab("Age") + ggtitle("Density distribution of age")
```

Exhibit 2: Divorce rate by age:

The chart below shows the divorce rate (as defined in the introduction) for each of the 83 age groups (18 to 100). 

The divorce rate for observations aged 18, 19, and 20 is significantly higher than observations in their 20's. The rationale for this observation is the scarcity (nominally) of married people who are 20 and less, which results in a large relative proportion of divorced people when compared to other age groups. The high divorce rate among young people can also be explained by the often painful and troubelsome reality of marriage. 
https://www.theguardian.com/lifeandstyle/2014/dec/19/-sp-why-do-so-many-young-marriages-come-to-an-early-end

The age with the highest divorce rate in the dataset is 48. Likely causes of this statistic are people who got married earlier in their lives and got divorced for reasons stated above, and now belong to this age group, or the shift in societal norms where staying in an un-happy marriage is no longer neccessary and where online dating and other activities are so readily available. 
http://www.prince-evans.co.uk/2017/01/divorce-rate-highest-among-aged-40-49/

This is further suported by the fact that younger people are less likely to be divorced, as those would be early-on in their marriages,while older people have a greater chance of being being remarried and thus will once again be classified as married according to the classifications in the UK census data. Additionally, although divorce rates for people above the age of 53 (Baby Boomers) are rising, this class of society generally has traditional values and beleif systems that consider marriage to be the norm. This, therefore, supports the idea that the marginal rate of divorce decreases above the age of 48. https://www.theguardian.com/lifeandstyle/2013/aug/06/divorces-rise-over-60s
  

```{r}

data2 <-  fd %>% filter(marstat == 2 | marstat == 3 | marstat == 4 | marstat == 5) %>% filter(age >= 18) %>% mutate(seperated = marstat) %>% mutate(seperated = replace(seperated, marstat!= 5 | marstat != 4, 0)) %>% mutate(seperated = replace(seperated, marstat == 5 | marstat == 4, 1))
#new dataframe containing the summarised count by age
total_group <- data2 %>% dplyr::select(age, seperated) %>% group_by(age) %>% summarise(count = n())
total_group_full <- total_group[1:52,]
dummy1 <- data.frame(age = 70:100)
dummy1$count <- rep(1,31)
dummy1$count[1:5] <- rep(round(total_group$count[53] / 5), 5)
dummy1$count[6:10] <- rep(round(total_group$count[54] / 5), 5)
dummy1$count[11:15] <- rep(round(total_group$count[55] / 5), 5)
dummy1$count[16:20] <- rep(round(total_group$count[56] / 5), 5)
dummy1$count[21:31] <- rep(round(total_group$count[57] / 11), 11)
new_total_group <- rbind.data.frame(total_group_full, dummy1)

#create a new sub-data frame from data2 grouped by age and seperated status and summarised 
divorced <-  data2 %>% dplyr::select(age, seperated) %>% filter(seperated == 1) %>% group_by(age, seperated) %>% summarise(n = n()) %>% ungroup() %>% group_by(seperated) %>% mutate(prop = n / sum(n) * 100)

#new divorced_full dataframe that contains all age groups above 70 i.e (70 to 100) since the orignal dataset groups ges above 70 by increments of 5. Count of divorced people was divided equally among the five tranches
divorced_full <- divorced[1:52,]
dummy <- data.frame(age = 70:100)
dummy$seperated <- rep(1,31)
dummy$n <- rep(1,31)
dummy$prop <- rep(1,31)
dummy$n[1:5] <- rep(divorced$n[53] / 5, 5)
dummy$prop[1:5] <- rep(divorced$prop[53] / 5, 5)
dummy$n[6:10] <- rep(divorced$n[54] / 5, 5)
dummy$prop[6:10] <- rep(divorced$prop[54] / 5, 5)
dummy$n[11:15] <- rep(divorced$n[55] / 5, 5)
dummy$prop[11:15] <- rep(divorced$prop[55] / 5, 5)
dummy$n[16:20] <- rep(divorced$n[56] / 5, 5)
dummy$prop[16:20] <- rep(divorced$prop[56] / 5, 5)
dummy$n[21:31] <- rep(divorced$n[57] / 11, 11)
dummy$prop[21:31] <- rep(divorced$prop[57] / 11, 11)
new_divorced <- rbind.data.frame(divorced_full, dummy)

#create new data.frame to calculate the divorce rate per age. This is done by dividing the count of divorced people from divorced data frame by the respective rows from the total group data frame 
divorce_by_age <- data.frame(prop = c(1:83), age = new_divorced$age)
divorce_by_age$prop <- rep(1, 83)
divorce_by_age$prop <- new_divorced$n / new_total_group$count * 100

ggplot(data = divorce_by_age, aes(x = age, y = prop)) + geom_line() + ylab("Divorce rate (%)") + xlab("Age") + ggtitle("Divorce rate by age")
```

Exhibit 3: Proportion of divorced people by age

The chart above shows the proportion of divorced people for each age group. The chart shows a curve resembling a normal distribution.
```{r}
ggplot(data = new_divorced, aes(x = age, y = prop)) + geom_line() + ylab("Divorce rate (%)") + xlab("Age") + ggtitle("Proportion of divorced people by age")
```

Exhibit 5: Divorce rate by age and social grade

The following chart depicts the distribution of divorce rate by age per level of social class.
From the data, it is apparent that the trend observed in the overall sample can also be observed when examining social classes, the divorce rate trends upwards from the lower ages and peaks in the late 40's, this trend then reverses with the divorce rate decreasing with age.

It may be worth noting that while the general trend is consistent accross social classes, one inconsistency does seem to appear in the 18 age group. The people aged 18 in the "AB" social class have a much lower divorce rate of around 8% than the other social classes at more than 20%. This seems to suggest that at young ages, affluence and wealth might potentially provide a cushion for a married couple and alleviate the stresses of financial troubles, which play a big role in divorce.

These findings provide support the notion that divorce rate distibution by age seems to behave similarly regardless of other variables. Further analysis may be required to assess the validity of this claim accross different variables.



```{r}
#creating a new dataframe grouped by age social grade and divorced status
data4 <- data2 %>% dplyr::select(age, scgpuk11c, seperated) %>% filter(scgpuk11c != -9) %>% filter(seperated == 1) %>% group_by(age, scgpuk11c, seperated) %>% summarise(n = n())
data4_control <- data2 %>% dplyr::select(age, scgpuk11c) %>% filter(scgpuk11c != -9) %>% group_by(age, scgpuk11c) %>% summarise(n = n())
data4$divorce_rate <- data4$n / data4_control$n * 100
data4$scgpuk11c <- factor(data4$scgpuk11c, levels = c(1,2,3,4), labels = c("AB", "C1", "C2", "DE"))
ggplot(data = data4, aes(x = age, y = divorce_rate)) + geom_col(position = 'dodge', colour = 'red') + facet_wrap(~scgpuk11c) + ylab("Divorce rate (%)") + xlab("Age") + ggtitle("Divorce rate distibution by age per social grade")

```

# Logistic Regression
A comprehensive regression analysis of all correlated variables

The variables presented thus far in this paper are the variables hypothesised to have the greatest effect on a person's likliness of being divorced in a certain society.

As the dependant variable in question is a dichotomous variable (divorced or not divorced), a logistic regression was performed where the divorced status is the dependant variable, and the variables mentioned above are the independant variables.

A logistic model is used to predict the probability of a binary response based on a set of other variables.

Regression formula:
divorced status = intercept + b1age + b2education_level + b3number_of_children + b4religion + b5london_or_not +...


As a measure of the overall model's explanatory powe, three main measures were use:
1) Akaike Information Criterion 
2) c-statistic or AUC (area under ROC curve)
3) The accuracy of the predictions above a certain threshold (the mean divorce rate for our purposes)

The best fit was determined by adding and removing the variables above until the best combination of the aforementioned metrics was obtained.

*insert model here*


# Conclusion

This paper has examined variables hypothesised to affect divorce rates in the UK. According to the analysis performed above, it is evident that divorce rates have multiple determinant factors, which themselves are potentially co-dependant. This analysis does not attempt to single out a specific variable or group of variables as the main determinant of divorce rate, but examines potential affecting variables.

Lacking any information about a particular person, age would be a good first predictor of marital status. It has been shown that people in their 40's have the highest divorce rates in society. This is true regardless of social status.

It is evident that social class and education level are amongst the biggest influencing factors of divorce rate as depicted by the strong inverse relationship of these variables with divorce rates. This trend holds true accross all ethnicities and religions, which further suppports the strength of the variables.




