---
title: "cen_fsus group project"
output: html_document
---


Over time and geographical space, it has been shown various times that divorce rate differs reagrding education social class and income. The higher educated and the higher the social class the lower the divorce rate. A the divorce rate in the UK has sunk by 34% in the last ten years, we want to evaluate if this purposes still hold (Office for National Statistics).

Social grade, a measure developped by the "National Readership Survey" should represent social class and income (It has been shown that social grade and income correlate highly). It is classified as follows: 

A	Higher managerial, administrative and professional	
B	Intermediate managerial, administrative and professional	
C1	Supervisory, clerical and junior managerial, administrative and professional	
C2	Skilled manual workers	
D	Semi-skilled and unskilled manual workers
E	State pensioners, casual and lowest grade workers, unemployed with state benefits only

In the original census data the social grades were grouped (AB, C1, C2, DE) and studends as well as people over 64 years old do not have a social grade and are therefore not included in the analysis.

The variable "Education" represents the highest degree people achieved. Holder of non-british awards were not classified in the original data and therefore not included in the analysis. 

The hypothesis underlying the following analysis claim a negative correlation between education and divorce rate, as well as a negative correlation between social class and divorce rate. 
Thanks to the 

```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(RCurl)
library(foreign)

url <- "https://raw.githubusercontent.com/dlouhasha/TheGreatWork/master/data/filtered_dataset.csv"
filtered_data <- getURL(url)
filtered_data <- read.csv(textConnection(filtered_data))

fd <- filtered_data
fd <- as.data.frame(fd)

census <- fd
View(census)

census$ethnicityew <- factor(census$ethnicityew, levels = seq(1,18), labels = c("English/Welsh/Scottish/Northern Irish/British", 
"Irish", 
"Gypsy or Irish Traveller", 
"Other White", 
"Mixed/multiple ethnic group: White and Black Caribbean", 
"Mixed/multiple ethnic group: White and Black African", 
"Mixed/multiple ethnic group: White and Asian", 
"Mixed/multiple ethnic group: Other Mixed", 
"Indian", 
"Pakistani", 
"Bangladeshi", 
"Chinese", 
"Other Asian", 
"African", 
"Caribbean", 
"Other Black", 
"Arab", 
"Any other ethnic group"))

#1: white 2: Asian 3: black 4: Arab 5: other

census$ethngrouped <- census$ethnicityew

census$ethngrouped <- ifelse(census$ethnicityew == "English/Welsh/Scottish/Northern Irish/British"  | census$ethnicityew == "Irish" | census$ethnicityew == "Gypsy or Irish Traveller"| census$ethnicityew == "Other White", 1, 
                                ifelse( census$ethnicityew == "Indian" | census$ethnicityew == "Pakistani"| census$ethnicityew == "Bangladeshi" | census$ethnicityew == "Chinese"| census$ethnicityew == "Other Asian", 2, 
                                        ifelse(census$ethnicityew == "African" | census$ethnicityew == "Caribbean"| census$ethnicityew == "Other Black", 3, 
                                               ifelse(census$ethnicityew == "Arab", 4, 
                                                      ifelse(census$ethnicityew == "other", "Any other ethnic group", 5)))))


census$religionew <- factor(census$religionew, levels = seq(1,9), labels = c("No religion", 
"Christian", 
"Buddhist", 
"Hindu", 
"Jewish", 
"Muslim", 
"Sikh", 
"Other religion", 
"Not stated"))

cen <- select(census, caseno, marstat, hlqupuk11, scgpuk11c)
cen$separated <- cen$marstat
cen <- cen %>% mutate(separated = replace(separated, marstat == 2 | marstat == 3, 0))
cen <- cen %>% mutate(separated = replace(separated, marstat == 5 | marstat == 4, 1))
colnames(cen) <- c("ID", "marstat", "education", "social_grade", "separated")
#replace -9 with NA 
cen_f <- cen %>% mutate_all(funs(replace(., . == -9, NA)))
cen_f <- cen_f %>% filter(!is.na(social_grade), !is.na(education), !(education == 16))
dim(cen_f)
```

After filtering out people whose education level cannot be interpretated and filtering out missing values 1.684.636 observations are left. 
The distribution between social grades is alike.
Education on the other hand has as significant underrepresentation of participants that absolved an apprenticeship. They account for less than 4 % of the surveyed participants . This should be kept in mind when interpreting divorcerates in the apprenticeship group.


Between education and social grade is a middle strong linear correlation. The higher the education level the higher the social grade.  Still it is insightful to look at both variables alone and combined, as they give different insights.
```{r}
#quick look at distributions
table(cen_f$separated)
addmargins(prop.table(table(cen_f$separated)))
table(cen_f$education)
prop.table(table(cen_f$education))
```


```{r}
#correlate independend variables 
library(ggplot2)
library(dplyr)
social_plot <- ggplot(cen_f, aes(x = factor(education), y = social_grade))
social_plot +
  geom_histogram(stat = "summary", fun.y = "mean", fill = "cadetblue3") +
  xlab ("Education level") +
  ylab("Average social Grade (1:highest 4: lowest)") +
  scale_x_discrete(labels = c("10" = "none", "11" = "O levels / GCSE", "12" = "CSE / GCSE grade 1 / A", "13" = "apprenticeship", "14" = "A levels", "15" = "university degree")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


cen_f %>% group_by(education) %>% summarise(n = n(), mean_grade = mean(social_grade, na.rm = TRUE))

#Spearmans ran correlation coefficient (cor between two comparative variables)
cor(cen_f$social_grade, cen_f$education, method = "spearman", use="p")

```



```{r, echo=FALSE}

#Plot divorce rate per social grade

educ <- cen_f %>% group_by(education) %>% summarise(n = n(), separated = sum(separated), rel_div = (separated / n)*100)

soclass <- cen_f %>% filter(!is.na(social_grade)) %>% group_by(social_grade) %>% summarise(n = n(), separated = sum(separated), rel_div = (separated / n)*100)

soclass_plot <- ggplot(soclass, aes(x = factor(social_grade), y = rel_div))
soclass_plot  +
  geom_col(fill  = "cadetblue3") +
  xlab("Social grade") +
  ylab("Divorce rate in percent") +
  scale_x_discrete(labels = c("1" = "AB", "2" = "C1", "3" = "C2", "4" = "DE")) +
  scale_y_continuous(limits = c(0, 33), breaks = seq(0, 33, 5))

```


For social grade the finding match the underlying hypothesis fundamentally. 
Social grade is directly connected to living standarts.Society holds the common belief that money is the root of all marriage problems. Nevertheless, there a few studies supporting this theory and many contradicting it (Andersen, 2008).
http://www.tandfonline.com/doi/abs/10.1300/J087v43n01_08  According to the UK census data, this assumption seems to prove true: the lowest social grade has an above average divorce rate of 33 % versus 13 % for the upper and middle class. Only between social grade C1 and C2 the pattern does not apply. C2 has a 4% lower divorce rate compared to C1. But still, both lie in the middle of AB and DE and are close to the total verage divorce rate of 21 %.

To see if the social grade is a true indicator for divorce and not only related to the fact, that migrants, different religions or higher educated, who tend to have a lower divorce rate than the Britishs, are overrepresented in some social groups, the data was crosschecked with religion as well as ethnicity. 

The exact same pattern, also with C1 being slightly higher than C2 holds across all religions, ethnicities and education levels. Therefore it can be conlcuded, that social grade has a direct influence on the divorcerate. Upper and middle class have a way lower probability of getting divorced than all other groups. 
A meaningful factor is the higher income, which enables better housing, better healthcare, less stressful life, more security and therefore less severe problems affecting the average marriage. 
Next to monetary problems, lower social groups tend to have a higher probability of drug and alcohol problem, one of the major reasons for divorce (NHS 2014; Trail and Karney, 2012). Additionally, earlier marriages, which will be explained in the education evaluation, are a burden for marriages.

http://digital.nhs.uk/catalogue/PUB15943
http://onlinelibrary.wiley.com/doi/10.1111/j.1741-3737.2012.00977.x/full

Across all sociodemographic characteristics (religion, ethnicity, education levels) the lowest social grade always has a significantly higher divorce rate and the highest social grade the lowest divorce rate. Overall it can be concluded that social grade, which is highly correlated with income, is a strong driver for divorces. 
The difference between social grade C1 and C2, which is as well constant across different sociodemographic groups, is a question for further research and could not be evaluated based on the census data. 


#EDUCATION
```{r}
#plot distribution of divorce per education level

#rel_div: percen_f_ft of divorce per education group
edu_plot <- ggplot(educ, aes(x = factor(education)))
edu_plot + 
  geom_col(aes(y = rel_div), fill = "cadetblue3") +
  geom_blank() +
  scale_x_discrete(labels = c("10" = "none", "11" = "O levels / GCSE", "12" = "CSE / GCSE grade 1 / A", "13" = "apprenticeship", "14" = "A levels", "15" = "university degree")) +
  scale_y_continuous(limits = c(0, 30), breaks = seq(0, 30, 3)) +
  ggtitle("Divorce rate per education level") +
  xlab ("Education level") +
  ylab ("Divore rate in %") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


The graph clearly confirms our expetations: the higher educated the lower the divorce rate. As can be seen, the difference in divorce rate between no degree and a university degree is about 10%. 
People with no education have a divorce rate of 28 %, univeristy graduates 17 %. People with a school education approach the mean divorce rate of 21 % the higher their school degree is. 

A reason for the negative correlation directly linked to education, is that higher educated people marry later than less educated people. A reason for  marrying later the higher educated someone is, is that the education lasts longer and therefore "adult life" and a stable lifestyle starts later. More educated people tend to focus more on their career in their joung ages.  Especially for women, a higher education results in a higher independence which lowers her feeling of need for getting married early (Isen, 2010).
Additionally, higher educated people are in a relatioship and live together longer time before getting married. Knowledge about the partner, expectations and security if both share the same ideas for the future grow with age  and experienced with each other.  https://www.theguardian.com/lifeandstyle/2009/sep/27/divorce-linked-to-age-education http://www.nber.org/papers/w15725.pdf
Furthermore, being pregnant or having children is still a reason for early marriage which results in a higher divorce rate. As less educated women tend to have children when their younger, they marry earlier because of the child (Office for National Statistics, 2015).

A result from the higher education is a better financial situation which lowerst the divorse rate, as seen in the social grade analysis. As 80% of the people marry in their own social category, the gap between income and education between different couples grow even stronger.

http://www.sueddeutsche.de/wirtschaft/partnerwahl-und-ungleichheit-heiraten-zementiert-soziale-spaltung-1.1875581 (http://www.pewresearch.org/fact-tank/2015/12/04/education-and-marriage/)

https://www.ons.gov.uk/peoplepopulationandcommunity/birthsdeathsandmarriages/livebirths/bulletins/birthsbyparentscharacteristicsinenglandandwales/2015

By looking at the relation between divorce rate and education across different religions and ethnic groups, it can be seen that the negative correlation holds. Education therefore seems to be a driver for divorce independend of the religion and ethnicity.
This strong pattern diminishes when looking at the highest degree in every social group. The negative correlation is still present, but as education is highly inequaly distributed among social grades, social grade acts as a confuncing variable.
The strong pattern of a lower divorce rate in the social group AB is partly driven by the majority of the people belonging to this social class being higly educated 

It can therefore be concluded that both, education and social grade, are strong driver of divorce.The unequally ditributed educaiton levels within social classes reinforce this effect. 
The highest overall divorce rate of 34% is thus in the lowest social grade group for people with no education at all. 
This effect is reinforced for black people in the lowest social grade with no education. Their divorce rate is 37% on average. With a divorce rate of only 6% highly educated Asians in the AB social group hold the lowest divorce rate. 

The average british person is white, christian, belongs to the middle class and has a university degree has a average divorce rate of 12 %. 
By only changing education (but keeping social class) the divorce rate ascends to 14.5 %.
When looking for both education extremes in the lowest social class we get a divorce rate for university graduates of 24 % versus 35 % for non educated people. It can be seen by this, that social grade accounts for higher differences in divorce rate than education does. 


```{r}
#grouping ethnicity
census$ethngrouped <- census$ethnicityew

census$ethngrouped <- ifelse(census$ethnicityew == "English/Welsh/Scottish/Northern Irish/British"  | census$ethnicityew == "Irish" | census$ethnicityew == "Gypsy or Irish Traveller"| census$ethnicityew == "Other White", 1, 
                                ifelse( census$ethnicityew == "Indian" | census$ethnicityew == "Pakistani"| census$ethnicityew == "Bangladeshi" | census$ethnicityew == "Chinese"| census$ethnicityew == "Other Asian", 2, 
                                        ifelse(census$ethnicityew == "African" | census$ethnicityew == "Caribbean"| census$ethnicityew == "Other Black", 3, 
                                               ifelse(census$ethnicityew == "Arab", 4, 
                                                      ifelse(census$ethnicityew == "other", "Any other ethnic group", 5)))))


educ <- cen_f %>% group_by(education, social_grade) %>% summarise(n = n(), separated = sum(separated), rel_div = (separated / n)*100) %>% mutate(sg = n/sum(n))

#add seperation per social grade 

edu_plot <- ggplot(educ, aes(x = factor(education), y = rel_div, fill = factor(social_grade)))
edu_plot + 
  geom_col(stat = "identity", position = "dodge") +
  scale_x_discrete(labels = c("10" = "none", "11" = "O levels / GCSE", "12" = "CSE / GCSE grade 1 / A", "13" = "apprenticeship", "14" = "A levels", "15" = "university degree")) +
  scale_y_continuous(limits = c(0, 35), breaks = seq(0, 35, 3)) +
  ggtitle("Divorce rate per education level") +
  xlab ("Education level") +
  ylab ("Divore rate in %") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#plot with facet_grid per social grade 
####
edu_plot <- ggplot(educ, aes(x = factor(education), y = rel_div, fill = factor(social_grade)))
edu_plot + 
  geom_col(stat = "identity") +
  scale_x_discrete(labels = c("10" = "none", "11" = "O levels / GCSE", "12" = "CSE / GCSE grade 1 / A", "13" = "apprenticeship", "14" = "A levels", "15" = "university degree")) +
  scale_y_continuous(limits = c(0, 35), breaks = seq(0, 35, 2.5)) +
  ggtitle("Divorce rate per education level and social grade") +
  xlab ("Education level") +
  ylab ("Divore rate in %") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(~social_grade)

#plot education per social grade, no divorce rate
edu_plot <- ggplot(educ, aes(x = factor(education), y =n, fill = factor(social_grade)))
edu_plot + 
  geom_col(stat = "identity") +
  scale_x_discrete(labels = c("10" = "none", "11" = "O levels / GCSE", "12" = "CSE / GCSE grade 1 / A", "13" = "apprenticeship", "14" = "A levels", "15" = "university degree")) +
  ggtitle("Divorce rate per education level and social grade") +
  xlab ("Education level") +
  ylab ("Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(~social_grade)
```




```{r}
educ <- cen_f %>% group_by(education, social_grade) %>% summarise(n = n(), separated = sum(separated), rel_div = (separated / n)*100) %>% mutate(sg = n/sum(n))

#add seperation per social grade 

edu_plot <- ggplot(educ, aes(x = factor(education), y = rel_div, fill = factor(social_grade)))
edu_plot + 
  geom_col(stat = "identity", position = "dodge") +
  scale_x_discrete(labels = c("10" = "none", "11" = "O levels / GCSE", "12" = "CSE / GCSE grade 1 / A", "13" = "apprenticeship", "14" = "A levels", "15" = "university degree")) +
  scale_y_continuous(limits = c(0, 35), breaks = seq(0, 35, 3)) +
  ggtitle("Divorce rate per education level") +
  xlab ("Education level") +
  ylab ("Divore rate in %") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#plot with facet_grid per social grade 
####
edu_plot <- ggplot(educ, aes(x = factor(education), y = rel_div, fill = factor(social_grade)))
edu_plot + 
  geom_col(stat = "identity") +
  scale_x_discrete(labels = c("10" = "none", "11" = "O levels / GCSE", "12" = "CSE / GCSE grade 1 / A", "13" = "apprenticeship", "14" = "A levels", "15" = "university degree")) +
  scale_y_continuous(limits = c(0, 35), breaks = seq(0, 35, 2.5)) +
  ggtitle("Divorce rate per education level and social grade") +
  xlab ("Education level") +
  ylab ("Divore rate in %") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(~social_grade)
```



